# Breeding nodes {#breeding}

Delineating breeding nodes is necessary for our migratory network model for both 1) assignment among populations, and 2) specifying a region for relative abundance. Here, we show how breeding nodes can be delineated by genetically distinct populations on the breeding grounds. In this example, we'll show how to use [eBird Status and Trends](https://ebird.org/science/status-and-trends) data to specify the breeding range and then use genetic data from admixture analyses to specify the spatial extent of the breeding nodes.

## ebirdst

In our migratory network analyses, the [eBird Status and Trends](https://science.ebird.org/en/status-and-trends) data is used to delineate the different stages of the annual cycle. Prior to doing anything with eBird Status and Trends data, you will need to download the `ebirdst` package, and then get access to the data. **You will need to follow the most up-to-date instructions from the `ebirdst` developers for getting the abundance data. Currently, that information is here: [https://ebird.github.io/ebirdst/](https://ebird.github.io/ebirdst/)**

To download the package:

``` r
# install.packages("remotes")
remotes::install_github("CornellLabofOrnithology/ebirdst")
```

Then, get access to `ebirdst` data at https://ebird.org/st/request. You will receive a key to download `ebirdst` data and you can enter that key in R:

``` r
ebirdst::set_ebirdst_access_key("XXXXX")
```

where `"XXXXX"` is the key.

By following instructions from the `ebirdst` developers, you can obtain polygons of the breeding and nonbreeding ranges of avian species (see [https://ebird.github.io/ebirdst/](https://ebird.github.io/ebirdst/) for details). 

## Creating the genoscape

A genoscape is the collection of genetically distinct populations that make up a species' range [@ruegg2021american]. Typically, for migratory species, the genoscape describes this population structure on the breeding range because the nonbreeding populations can can contain individuals from different breeding populations.

We will outline the main genoscape creation steps here, but full instructions on creating a genoscape map can be found in Eric Anderson's Github project [Make a Bird Genoscape Project map](https://github.com/eriqande/make-a-BGP-map). The input data needed for a genoscape are:

* Individual Q-value matrix
* Lat/lon matrix of individuals
* Breeding range polygon

The Q-value matrix is obtained from individual admixture analyses (e.g. Structure, Admixture, `snmf` fuction from the LEA R-package). Latitude/longitude coordinates are for the individual samples used in the Q-value matrix. Breeding range polygons can be obtained from `ebirdst` (see previous section). 

The `comyel_breeding_data` data set provides admixture results (Q-values) of five genotype clusters for Common Yellowthroat (cite a coye paper) and metadata for the sampled individuals.

``` r
Q_matrix <- Mignette::comyel_breeding_data %>%
  dplyr::select(CA, Midwest, NewEngland, West, Southwest) %>%
  as.matrix()
long_lat_matrix <- Mignette::comyel_breeding_data %>%
  dplyr::select(Long, Lat) %>%
  as.matrix()
cluster_colors <-  c(
  CA = "#CC0000",
  Midwest = "#3399FF",
  NewEngland = "#9933CC",
  West = "#009933",
  Southwest = "#FF6600") 
```

We will use a [modified version](https://github.com/eriqande/TESS3_encho_sen) of the `tess3r` package to create the genoscape rasters.

``` r
# remotes::install_github("eriqande/TESS3_encho_sen")
genoscape_brick <- tess3r::tess3Q_map_rasters(
  x = Q_matrix, 
  coord = long_lat_matrix,  
  map.polygon = breed_smooth,
  window = terra::ext(breed_smooth)[1:4],
  resolution = c(300,300), # if you want more cells in your raster, set higher
  col.palette = tess3r::CreatePalette(cluster_colors, length(cluster_colors)), 
  method = "map.max", 
  interpol = tess3r::FieldsKrigModel(10),  
  main = "Ancestry coefficients",
  xlab = "Longitude", 
  ylab = "Latitude", 
  cex = .4
)
names(genoscape_brick) <- colnames(Q_matrix)

out.files <- paste0("./comyel/genoscape/comyel_genoscape_cluster_", names(genoscape_brick), ".tif")
terra::writeRaster(terra::rast(genoscape_brick), filename = out.files)
```

**STOP**

The rasters for the genoscape are all that are needed for obtaining information on relative abundance for the different populations. You can continue on to the [relative abundance section](#abundance) if you are ready to do that with the genoscape. Or if you still need to create the wintering nodes, check out the [wintering nodes section](#wintering). The following section is not necessary for the migratory network but details how to covert genoscape rasters to polygons if the `mignette` user is interested in doing so.

### Genoscape polygons

Using the genoscape rasters we will convert them to polygons, using the handy `scape_to_shape()` function. The `prob_threshold` parameter specifies the value to determine if a raster cell is included in the polygon for that genoscape. This value should be customized for different species to check for overlap of genoscape polygons, which is not desirable. Setting too high of a threshold will create very small breeding nodes, while too low of a threshold will result in large, overlapping breeding nodes.

``` r
genoscape_files <- list.files("./comyel/genoscape", 
                                     full.names = T,
                                     pattern = "*.tif")

genoscape_raster_stack <- terra::rast(genoscape_files)
genoscape_polygon_sf <- mignette::scape_to_shape(x = genoscape_raster_stack, prob_threshold = 0.5)
```

Check out the polygons

```{r, echo = F, results = F, message = F}
library(sf)
library(ggplot2)
library(mignette)
genoscape_polygon_sf <- st_read(dsn = "./data/comyel/genoscape/polygons", 
                                layer = "comyel_genoscape_polygon_sf")
cluster_colors <-  c(
  CA = "#CC0000",
  Midwest = "#3399FF",
  NewEngland = "#9933CC",
  West = "#009933",
  Southwest = "#FF6600") 
```

```{r, out.width='70%'}
ggplot() +
  geom_sf(data = genoscape_polygon_sf,alpha = 0.75, aes(fill = Cluster)) +
  scale_fill_manual(values = cluster_colors)
```

Check which polygons are overlapping. Each row of the output provides a pair of overlapping polygons (if there are any).

```{r}
mignette::check_genoscape_overlap(genoscape_polygon_sf)
```

Now on to creating the [wintering nodes](#wintering)
