# Step 1) Define breeding and nonbreeding nodes {#nodes}

The first step of creating a migratory network is to spatially define the nodes (i.e. populations) that make up the sampled portions of the breeding and nonbreeding range. This can either be done with *vector* data for discrete spatial boundaries of nodes or *raster* data for a continuous surface probabilistic membership of geographic cells to a node (e.g., population genetic structure of a genoscape [@ruegg2020genoscape]).

## Vector data

Vector data can define nodes by representing geographic boundaries such as political borders or biogeographic regions. In `mignette`, we provide data for *conservation regions* of the Western Hemisphere that integrate geopolitical boundaries and ecoregions.

```{r}
# SpatVector from terra package
regions_file <- system.file("extdata", "conservation_regions.Rds", package = "mignette")
# read in file as SpatVector
conservation_regions <- terra::vect(regions_file)

terra::plot(conservation_regions, col=rainbow(10))
```

For defining a node, we are only interested in regions that we have data. It's straightforward with the spatial functions from `terra` to determine the sampled regions. Here, we find the conservation regions for which we have sampled American Redstart data from the nonbreeding range.

```{r}
# create SpatVector points object from Redstart lat/lon sampling data
nonbreeding_coords <- terra::vect(mignette::amre_nonbreeding_data, 
                             geom=c("Lon", "Lat"),
                             crs = "EPSG:4326")
# find spatial intersection of points with conservation regions
sampled_conservation_regions <- terra::intersect(nonbreeding_coords,
                                        conservation_regions)
# count the number of sampled individuals associated with each conservation region
sampled_table <- table(sampled_conservation_regions$Region)
sampled_table
```

Downstream in the workflow, we need sufficiently sampling of individuals in (or inference to) a node. So, we'll only retain the conservation regions with more than 3 samples.

```{r}
# retain conservation region names with more than 3 samples
retained_IDs <- names(sampled_table[sampled_table > 3])
retained_IDs
```

Thus, we've identified the 5 nonbreeding nodes for the American Redstart data: Amazon/Orinoco-Northern Uplands (AONU), Atlantic Lowland Mexico (ALM), Caribbean (CAR), Highland Central America (HCA), and Lowland Central America (LCA). We can plot these conservation regions along with the corresponding sampling points.

```{r}
# subset conservation regions file by retained regions
conservation_regions_subset <- terra::subset(conservation_regions, 
                                             conservation_regions$Region %in% retained_IDs)
# subset sampled points by retained regions
nonbreeding_coords_subset <- terra::subset(sampled_conservation_regions,
                                           sampled_conservation_regions$Region %in% retained_IDs)
# plot
terra::plot(conservation_regions_subset, col = rainbow(5))
terra::points(nonbreeding_coords_subset, cex = 1.5)
```


## Raster data

Raster data is used when populations are defined by a continuous surface of probabilities of membership to a population. In the example here, we have population genetic data from the American Redstart that define breeding populations (i.e., American Redstart genoscape, [@desaix2023low]). We have defined these nodes as Western Boreal (WB), Basin Rockies (BR), Northern Temperate (NT), Southern Temperate (ST) and Maritime Provinces (MP). Raster data for nodes are used in `mignette` to get node abundance that is scaled by the raster probabilities of population membership, as shown in the next section.


```{r, results = F, message = F}
amre_genoscape_file <- system.file("extdata", "amre_genoscape.Rds", package = "mignette")
amre_genoscape <- terra::rast(amre_genoscape_file)
terra::plot(amre_genoscape)
```

While it is outside the scope of `mignette` to perform the requisite population genetics analyses for producing these rasters, users with genetic data and expertise with genetic clustering analyses may be interested in the [Supplementary information](#supplemental) that provide additional information and code.

Now on to [Step 2) Abundance and migratory connectivity data](#data)
