# Step 3) Modeling the migratory network {#model}

## Input data

As described in the previous section the data required are:

* Abundance
* Migratory connectivity

The abundance data need to be in the following format, with node IDs (same names as in the *assignment* file) in the first column and abundance values in the second column.

```{r, echo=FALSE}
knitr::kable(mignette::amre_abundance)
```

For the migratory connectivity data, the input data needs to correspond to the following format:

```{r, echo=FALSE}
knitr::kable(mignette::amre_assign)
```

If you skipped it, see the previous [Step 2) Abundance and migratory connectivity data](#data) section for specific details on the input data formatting!

For the following functions, we specify the order of the nodes we are using for the model. Here, we are just ordering nodes geographically by longitude to facilitate straightforward interpretation of the output.

```{r}
bnode_names <- c("WB", "BR", "NT", "ST", "MP")
wnode_names <- c("ALM", "LCA", "HCA", "CAR", "AONU")
```


For the American Redstart migratory network, we use `model = 1` from `mignette` which specifies that nonbreeding nodes are “encountered” and breeding nodes are “recovered” (i.e., inferred). This output saves the model as `amre.genetic.model_1.txt`. Below we specify `parallel = TRUE` to run MCMC on multiple cores and use the remaining defaults described previously CHANGE LINK. This step is computationally intensive and takes ~2 minutes to run on a 2023 MacBook Pro with an Apple M2 Pro chip.


```r
network_model <- run_network_model(abundance = mignette::amre_abundance,
                       	nb2br_assign = mignette::amre_assign,
                       	bnode_names = bnode_names,
                       	wnode_names = wnode_names,
                       	model = 1, 
                       	base_filename = "amre.genetic",
                       	parallel = TRUE)
```

The `run_network_model()` function outputs a list object with two components. The first component of the output is `network_model$conn` which provides a formatted data frame (i.e., breeding nodes in first column, remaining columns have nonbreeding node names) of the modeled connectivity results. Each value in the connectivity matrix indicates the overall importance of a particular migratory connection. These values are interpreted as the proportion of individuals that migrate between the corresponding populations, as such all of the values in the connectivity matrix sum to one. 

```r
network_model$conn
```

```{r, echo=FALSE}
conn_ex <- mignette::amre_conn
rownames(conn_ex) <- NULL
conn_ex <- tibble::as_tibble(conn_ex) 
conn_ex <- tibble::add_column(conn_ex, "Breeding" = bnode_names, .before = 1)
knitr::kable(conn_ex)
```


The second component is the full output from [`*jagsUI* autojags()`](https://rdrr.io/cran/jagsUI/man/autojags.html) and is accessed by `network_model$jags_out`. Here, the raw connectivity matrix can be accessed:

```r
amre_conn <- network_model$jags_out$mean$conn_g
amre_conn
```

```{r, echo = FALSE}
amre_conn <- mignette::amre_conn
rownames(amre_conn) <- NULL
colnames(amre_conn) <- NULL
knitr::kable(amre_conn)
```

The raw connectivity matrix is used to plot the network. We plot the migratory network  with the provided `mignette` functions `net_create()` and `net_draw()`. We set `connected_tol = 0.01` which plots only the edges with connectivity values of greater than 0.01. We also demonstrate below how to change parameters such as the colors and the node and edge scale size.

```{r}
amre_net <- mignette::net_create(amre_conn,
                        	node.names = list(bnode_names,wnode_names),
                        	connected_tol = 0.02)
#set the display size range for nodes (min and max), default 1-10
amre_net$display_par$node_size_scale <- c(8,25)
#set the display size range for edges (min and max), default 1-10
amre_net$display_par$edge_size_scale <- c(1,5)
# change colors
amre_net$display_par$bnode_colors <- c("#009e73", "#cc79a7", "#56b4e9", "#e69f00", "#7979ff")
amre_net$display_par$wnode_colors <- "grey80"

mignette::net_draw(amre_net)
```

In this visualization, node size corresponds to the amount of connectivity with that population and edge size corresponds to the amount of connectivity between the populations. Breeding populations are in the top row, for which we provided custom colors, and nonbreeding populations are in the bottom row.

Now you have a migratory network! Check out the [visualization supplement](#visualization) for additional ideas on plotting the network.
